MIMO Channel Shaping and Rate Maximization\\Using Beyond Diagonal RIS

### 2025-02-13 manuscript/main.tex
:construction: Fix plot legend

\begin{abstract}
	This paper investigates the limits to which a passive \gls{ris} can shape a point-to-point \gls{mimo} in terms of singular values
	% and power gain
	and thus improve the rate performance.
	% redistribute the singular values of a point-to-point \gls{mimo} channel for providing power and rate gains.
	We depart from the conventional \gls{ris} with \gls{d} phase shift matrix and adopt a \gls{bd} model interconnecting elements for wave amplitude and phase manipulation.
	This translates to superior branch matching and mode alignment capabilities in \gls{mimo}.
	% featuring group-wise connections for enhanced branch matching and mode alignment capabilities.
	Specifically, analytical singular values bounds are derived under popular \gls{ris} deployment scenarios
	% the scenario of rank-deficient backward/forward channel or negligible direct channel, which
	to showcase the shaping advantages of \gls{bd}-\gls{ris} in terms of dynamic range and \gls{dof}.
	A numerical optimization method is then proposed for a broader range of shaping problems and invoked to characterize the entire achievable singular value region.
	As a side product, we tackle \gls{bd}-\gls{ris}-aided \gls{mimo} rate maximization problem by a local-optimal \gls{ao} and a low-complexity shaping-inspired approach.
	Results show that compared to \gls{d}-\gls{ris}, \gls{bd}-\gls{ris} significantly improves the dynamic range of all channel singular values and the trade-off in manipulating them, resulting in enhanced power gain and achievable rate.
	Those benefits become more pronounced when the number of \gls{ris} elements or \gls{mimo} dimensions increase.
	Of particular interest, \gls{bd}-\gls{ris} is shown to activate multi-stream transmission (hence achieving the asymptotic \gls{dof}) at much lower transmit power than \gls{d}-\gls{ris} thanks to its singular value shaping proficiency.
\end{abstract}

@@

	% \begin{subsection}{Related Works}
		At a specific resource block, channel shaping metrics can be classified into two categories.
		\begin{itemize}
			\item \emph{Singular value:} The impact of \gls{ris} has been studied in terms of minimum singular value \cite{ElMossallamy2021}, effective rank \cite{ElMossallamy2021,Meng2023}, condition number \cite{Zheng2022,Huang2023}, and degree of freedom \cite{Bafghi2022,Zheng2023,Chae2023}. Those are closely related to performance measures (e.g., achievable rate and harvested power \cite{Shen2021}) but sensitive to minor perturbations of the channel matrix.
			\item \emph{Power:} The impact of \gls{ris} has been studied in terms of channel power gain \cite{Wu2019,Shen2020a,Nerini2023,Nerini2024,Santamaria2023} in point-to-point channels and leakage interference \cite{Santamaria2023a} in interference channels. Those second-order metrics are less informative in \gls{mimo} but easier to analyze and optimize.
		\end{itemize}

		Although above works offered inspiring glimpses into the channel shaping potential of passive \gls{ris}, none attempted to disclose the entire singular value region.
		Most relevant literature \cite{ElMossallamy2021,Meng2023,Zheng2022,Huang2023,Bafghi2022,Zheng2023,Chae2023,Wu2019,Santamaria2023a} have also been limited to a \gls{d}-\gls{ris} model where each element is connected to a dedicated impedance and disconnected from others.
		As such, wave impinging on one element is entirely scattered by the same element.
		This simple architecture is modeled by a diagonal scattering matrix with unit-magnitude diagonal entries, which only applies a phase shift to the incoming signal.
		The idea was soon generalized to \gls{bd}-\gls{ris} with group-connected architecture \cite{Shen2020a}, where adjacent elements within the same group are connected via passive reconfigurable components\footnote{Those components can be either symmetric (e.g., capacitors and inductors) or asymmetric (e.g., ring hybrids and branch-line hybrids) \cite{Ahn2006}, resulting in symmetric and asymmetric scattering matrices, respectively.}.
		This allows wave impinging on one element to propagate within the circuit and depart partially from any element in the same group.
		It can thus manipulate both amplitude and phase of the scattered wave while remaining passive.
		Such a powerful model can be realized at reduced hardware cost using tree- and forest-connected architectures by graph theory \cite{Nerini2024}.
		\gls{bd}-\gls{ris} can also function in
		% hybrid transmitting-and-reflecting mode \cite{Li2023b} and
		multi-sector mode \cite{Li2023c} for full-space coverage and multi-user support.
		Practical challenges such as channel estimation \cite{Li2024}, mutual coupling \cite{Li2023f}, and wideband modelling \cite{Li2024a} have also been studied in recent literature.
		Its beamforming effectiveness over \gls{d}-\gls{ris} and energy efficiency over active \gls{ris} and relay systems have been proved in \gls{siso} and \gls{miso} systems \cite{Shen2020a,Nerini2023,Nerini2024,Santamaria2023,Fang2023,Zhou2023,Li2023c,Soleymani2024}.
		However, the interplay between \gls{bd}-\gls{ris} and \gls{mimo} is still in the infancy stage.
		The authors of \cite{Bartoli2023} investigated the rate-optimal joint beamforming design for a fully-connected \gls{bd}-\gls{ris}-aided \gls{mimo} system where the direct channel is negligible.
		A transmitter-side \gls{bd}-\gls{ris} was introduced to
		massive \gls{mimo} systems that exploits statistical \gls{csi} for improved spectral efficiency \cite{Mishra2024}, which again assumed negligible direct channel and fully-connected \gls{bd}-\gls{ris}.
		Received power maximization with continuous-valued and discrete-valued \gls{bd}-\gls{ris} have been tackled respectively in closed form \cite{Nerini2023} and by machine learning approach \cite{Nerini2023b}, but the corresponding single-stream transceiver is rate-suboptimal.
	% \end{subsection}

@@

		Fifth, we conduct extensive simulations to validate the analytical bounds and numerical methods.
		It is concluded that:
		\begin{itemize}
			\item \gls{bd}-\gls{ris} significantly improves the dynamic range of all channel singular values and the trade-off in manipulating them, resulting in enhanced power gain and achievable rate;
			\item The shaping advantages of \gls{bd}-\gls{ris} increase with the number of scattering elements and \gls{mimo} dimensions;
			\item \gls{bd}-\gls{ris} can activate multi-stream transmission (hence achieving the asymptotic \gls{dof}) at low transmit power thanks to its singular value shaping proficiency;
			\item The rate gap between the optimal and low-complexity beamforming designs diminishes as the \gls{ris} evolves from \gls{d} to fully-connected \gls{bd};
			\item The additional optimization cost of \gls{bd}-\gls{ris} over \gls{d}-\gls{ris} is affordable and the geodesic \gls{rcg} method is efficient on large-scale problems;
			\item All proposed asymmetric \gls{bd}-\gls{ris} designs can be modified for symmetry with reasonable performance cost.
		\end{itemize}
	% \end{subsection}

@@

\begin{section}{System Model}
	% \begin{subsection}{BD-RIS}
		The \gls{bd}-\gls{ris} can be modeled as an $N_\mathrm{S}$-port network \cite{Ivrlac2010} that divides into $G$ individual groups, where group $g \in [G]$ contains $N_g$ scattering elements interconnected by real-time reconfigurable components \cite{Shen2020a}.
		Apparently $N_\mathrm{S} = \sum_{g=1}^G N_g$.
		Without loss of generality we assume equal size for all groups $N_g = L \triangleq N_\mathrm{S} / G, \ \forall g$ and no mutual coupling between elements.
		For asymmetric \gls{bd}-\gls{ris}, the overall scattering matrix is block-diagonal with unitary blocks\footnote{We assume asymmetric network by default to establish a benchmark for passive \gls{ris}. Some symmetric solutions (i.e., $\mathbf{\Theta} = \mathbf{\Theta}^\mathsf{T}$) will be discussed and their performance will be evaluated in Section~\ref{sc:simulation}.}
		\begin{equation}
			\label{eq:ris}
			\mathbf{\Theta} = \diag(\mathbf{\Theta}_1,\ldots,\mathbf{\Theta}_G),
		\end{equation}
		where $\mathbf{\Theta}_g \in \mathbb{U}^{L \times L}$ is the $g$-th diagonal block modeling the response of group $g$.
		It is noteworthy that \gls{d}-\gls{ris} is an extreme case of \eqref{eq:ris} with group size $L=1$, that is, each element is an individual.
		Some viable architectures of \gls{bd}-\gls{ris} are illustrated in \cite[Fig. 3]{Shen2020a}, \cite[Fig. 5]{Li2023c}, \cite[Fig. 2]{Nerini2024} where the array geometry and circuit topology are modeled in $\mathbf{\Theta}$.
	% \end{subsection}

	% \begin{subsection}{MIMO Point-to-Point Channel}
		Consider a \gls{bd}-\gls{ris} aided \gls{mimo} \gls{pc} with $N_\mathrm{T}$ and $N_\mathrm{R}$ transmit and receive antennas, respectively, and $N_\mathrm{S}$ scattering elements at the \gls{bd}-\gls{ris}.
		This configuration is denoted as $N_\mathrm{T} \times N_\mathrm{S} \times N_\mathrm{R}$.
		Let $\mathbf{H}_\mathrm{D} \in \mathbb{C}^{N_\mathrm{R} \times N_\mathrm{T}}$, $\mathbf{H}_\mathrm{B} \in \mathbb{C}^{N_\mathrm{R} \times N_\mathrm{S}}$, $\mathbf{H}_\mathrm{F} \in \mathbb{C}^{N_\mathrm{S} \times N_\mathrm{T}}$ denote the direct (i.e., transmitter-receiver), backward (i.e., \gls{ris}-receiver), and forward (i.e., transmitter-\gls{ris}) channels, respectively.
		The equivalent channel depends on the \gls{bd}-\gls{ris} scattering matrix
		\begin{equation}
			\label{eq:channel}
			\mathbf{H} = \mathbf{H}_\mathrm{D} + \mathbf{H}_\mathrm{B} \mathbf{\Theta} \mathbf{H}_\mathrm{F} = \mathbf{H}_\mathrm{D} + \sum_g \mathbf{H}_{\mathrm{B},g} \mathbf{\Theta}_g \mathbf{H}_{\mathrm{F},g},
			%  \triangleq \mathbf{H}_\mathrm{D} + \sum_g \mathbf{H}_g,
		\end{equation}
		where $\mathbf{H}_{\mathrm{B},g} \in \mathbb{C}^{N_\mathrm{R} \times L}$ and $\mathbf{H}_{\mathrm{F},g} \in \mathbb{C}^{L \times N_\mathrm{T}}$ are the backward and forward channels associated with group $g$, corresponding to the $(g{-}1)L{+}1$ to $gL$ columns of $\mathbf{H}_\mathrm{B}$ and rows of $\mathbf{H}_\mathrm{F}$, respectively.
		% , and $\mathbf{H}_g \triangleq \mathbf{H}_{\mathrm{B},g} \mathbf{\Theta}_g \mathbf{H}_{\mathrm{F},g}$ is the indirect channel via group $g$.
		Since unitary matrices constitute an algebraic group with respect to multiplication, the scattering matrix of group $g$ can be decomposed as
		\begin{equation}
			\mathbf{\Theta}_g = \mathbf{L}_g \mathbf{X}_g \mathbf{R}_g^\mathsf{H},
		\end{equation}
		where $\mathbf{L}_g, \mathbf{R}_g \in \mathbb{U}^{L \times L}$ are two unitary matrices and $\mathbf{X}_g \in \mathbb{R}^{L \times L}$ is a permutation matrix.
		Let $\mathbf{H}_g \triangleq \mathbf{H}_{\mathrm{B},g} \mathbf{\Theta}_g \mathbf{H}_{\mathrm{F},g}$ be the indirect channel via group $g$ and $\mathbf{H}_{\mathrm{B/F},g} = \mathbf{U}_{\mathrm{B/F},g} \mathbf{\Sigma}_{\mathrm{B/F},g} \mathbf{V}_{\mathrm{B/F},g}^\mathsf{H}$ be the \gls{svd} of the backward and forward channels, respectively.
		The equivalent channel is thus
		\begin{equation}
			\label{eq:channel_svd}
			\mathbf{H} = \overbrace{\mathbf{H}_\mathrm{D} + \sum_g \mathbf{U}_{\mathrm{B},g} \mathbf{\Sigma}_{\mathrm{B},g} \underbrace{\mathbf{V}_{\mathrm{B},g}^\mathsf{H} \mathbf{L}_g \mathbf{X}_g \mathbf{R}_g^\mathsf{H} \mathbf{U}_{\mathrm{F},g}}_\text{backward-forward} \mathbf{\Sigma}_{\mathrm{F},g} \mathbf{V}_{\mathrm{F},g}^\mathsf{H}}^\text{direct-indirect}.
		\end{equation}

		\begin{remark}
			In \eqref{eq:channel_svd}, the \gls{bd}-\gls{ris} performs a blockwise unitary transformation to combine the backward-forward (intra-group, multiplicative) channels and direct-indirect (inter-group, additive) channels.
			These two attributes are refined as:
			\begin{itemize}
				\item \emph{Branch matching:} It refers to pairing and combining the branches (i.e., entries) of $\mathbf{H}_{\mathrm{B},g}$ and $\mathbf{H}_{\mathrm{F},g}$ through $\mathbf{\Theta}_g$.
				\item \emph{Mode alignment:} It refers to aligning and ordering the modes (i.e., singular vectors) of $\{\mathbf{H}_g\}_{g \in [G]}$ with those of $\mathbf{H}_\mathrm{D}$ through $\{\mathbf{\Theta}_g\}_{g \in [G]}$.
			\end{itemize}
		\end{remark}

		\begin{example}[\gls{siso} channel gain maximization]
			\label{eg:siso}
			Denote the direct, backward, forward channels as $h_\mathrm{D}$, $\mathbf{h}_\mathrm{B} \in \mathbb{C}^{N_\mathrm{S} \times 1}$, and $\mathbf{h}_\mathrm{F}^\mathsf{H} \in \mathbb{C}^{1 \times N_\mathrm{S}}$, respectively.
			In this case, mode alignment boils down to phase matching and the optimal \gls{bd}-\gls{ris} structure is
			% any $L \in [N_\mathrm{S}]$, including \gls{d}-\gls{ris}, suffices for perfect mode alignment using
			\begin{equation}
				\mathbf{\Theta}_{\textnormal{P-max},g}^\textnormal{SISO} = \frac{h_\mathrm{D}}{\lvert h_\mathrm{D} \rvert} \mathbf{V}_{\mathrm{B},g} \mathbf{U}_{\mathrm{F},g}^\mathsf{H}, \quad \forall g,
			\end{equation}
			where $\mathbf{V}_{\mathrm{B},g} = \bigl[\mathbf{h}_{\mathrm{B},g}/\lVert \mathbf{h}_{\mathrm{B},g} \rVert, \mathbf{N}_{\mathrm{B},g}\bigr] \in \mathbb{U}^{L \times L}$, $\mathbf{U}_{\mathrm{F},g} = \bigl[\mathbf{h}_{\mathrm{F},g}/\lVert \mathbf{h}_{\mathrm{F},g} \rVert, \mathbf{N}_{\mathrm{F},g}\bigr] \in \mathbb{U}^{L \times L}$, and $\mathbf{N}_{\mathrm{B/F},g} \in \mathbb{C}^{L \times (L-1)}$ are the orthonormal bases of kernels of $\mathbf{h}_{\mathrm{B/F},g}$.
			Apparently, any group size $L$ (including \gls{d}-\gls{ris}) suffices for perfect phase matching.
			The resulting maximum channel gain is a function of $L$
			\begin{equation}
				\ \lvert h \rvert = \lvert h_\mathrm{D} \rvert + \sum_g \sum_l \lvert h_{\mathrm{B},g,\pi_{\mathrm{B},g}(l)} \rvert \lvert h_{\mathrm{F},g,\pi_{\mathrm{F},g}(l)} \rvert,
			\end{equation}
			where $h_{\mathrm{B/F},g,l}$ are the $l$-th entries of $\mathbf{h}_{\mathrm{B/F},g}$, and $\pi_{\mathrm{B/F},g}$ are permutations of $[L]$ sorting their magnitude in similar orders.
			That is, the maximum \gls{siso} channel gain is attained when each \gls{bd}-\gls{ris} group, apart from phase shifting, matches the $l$-th strongest backward and forward channel branches.
			A larger $L$ provides more flexible branch matching and thus higher channel gain.
		\end{example}

@@

	\begin{subsection}{Analytical Shaping Bounds}
		\label{sc:shaping_analytical}
		The main results of this subsection are presented in the following Propositions and Corollaries.
		\begin{proposition}[Degrees of freedom]
			\label{pp:dof}
			\gls{bd}-\gls{ris} may achieve a larger or smaller \gls{mimo} \gls{dof}\footnote{Here communication \gls{dof} refers to the maximum number of independent streams that can be transmitted in parallel over a \gls{mimo} channel, i.e., $\mathrm{DoF}(\mathbf{H}) = \lim_{\rho \to \infty} \frac{\log \det(\mathbf{I} + \rho \mathbf{H} \mathbf{H}^\mathsf{H})}{\log \rho}$ where $\rho$ is the \gls{snr}. The concept is closely related to radar \gls{dof} which refers to the number of independent measurements.} than \gls{d}-\gls{ris}.
		\end{proposition}
		\begin{proof}
			Please refer to Appendix~\ref{ap:dof}.
		\end{proof}

		While increasing the \gls{dof} improves the asymptotic rate performance for point-to-point transmission, the ability to reduce the \gls{dof} can be exploited to orthogonalize channels and align interference in multi-user scenarios.
		% reducing the \gls{dof} does not necessarily degrade the rate at finite \gls{snr}.
		% , it does not necessarily imply a higher rate at finite \gls{snr}.
		% It is worth mentioning that a larger \gls{dof} implies more parallel streams or less interference, which is beneficial for high-rate communication.

>>>>

\begin{abstract}
	This paper investigates the limits to which a passive \gls{ris} can reshape a point-to-point \gls{mimo} in terms of singular values
	% and power gain
	and thus improve the rate performance.
	% redistribute the singular values of a point-to-point \gls{mimo} channel for providing power and rate gains.
	We depart from the conventional \gls{ris} with \gls{d} phase shift matrix and adopt a \gls{bd} model interconnecting elements for signal amplitude and phase manipulation.
	This translates to superior branch matching and mode alignment capabilities in \gls{mimo}.
	% featuring group-wise connections for enhanced branch matching and mode alignment capabilities.
	Specifically, analytical singular values bounds are derived under popular \gls{ris} deployment scenarios
	% the scenario of rank-deficient backward/forward channel or negligible direct channel, which
	to showcase the shaping advantages of \gls{bd}-\gls{ris} in terms of dynamic range and \gls{dof}.
	A numerical optimization method is then proposed for a broader range of shaping problems and invoked to characterize the achievable singular value region.
	As a side product, we tackle \gls{bd}-\gls{ris}-aided \gls{mimo} rate maximization problem by a local-optimal \gls{ao} and a low-complexity shaping-inspired approach.
	Results show that compared to \gls{d}-\gls{ris}, \gls{bd}-\gls{ris} significantly improves the dynamic range of all channel singular values and the trade-off in manipulating them, resulting in enhanced power gain and achievable rate.
	Those benefits become more pronounced when the number of \gls{ris} elements or \gls{mimo} dimensions increase.
	Of particular interest, \gls{bd}-\gls{ris} is shown to activate multi-stream transmission (hence achieving the asymptotic \gls{dof}) at much lower transmit power than \gls{d}-\gls{ris} thanks to its singular value shaping proficiency.
\end{abstract}

@@

	% \begin{subsection}{Related Works}
		At a specific resource block, channel shaping metrics can be classified into two categories.
		\begin{itemize}
			\item \emph{Singular value:} The impact of \gls{ris} has been studied in terms of minimum singular value \cite{ElMossallamy2021}, effective rank \cite{ElMossallamy2021,Meng2023}, condition number \cite{Zheng2022,Huang2023}, and degree of freedom \cite{Bafghi2022,Zheng2023,Chae2023}. Those are closely related to performance measures (e.g., achievable rate and harvested power \cite{Shen2021}) but sensitive to minor perturbations of the channel matrix;
			\item \emph{Power:} The impact of \gls{ris} has been studied in terms of channel power gain \cite{Wu2019,Shen2020a,Nerini2023,Nerini2024,Santamaria2023} in point-to-point channels and leakage interference \cite{Santamaria2023a} in interference channels. Those second-order metrics are less informative in \gls{mimo} but easier to analyze and optimize.
		\end{itemize}

		Although above works offered inspiring glimpses into the channel shaping potential of passive \gls{ris}, none attempted to disclose the entire singular value region.
		Most relevant literature \cite{ElMossallamy2021,Meng2023,Zheng2022,Huang2023,Bafghi2022,Zheng2023,Chae2023,Wu2019,Santamaria2023a} have also been limited to a \gls{d}-\gls{ris} model where each element is connected to a dedicated impedance and disconnected from others.
		As such, wave impinging on one element is entirely scattered by the same element.
		This simple architecture is modeled by a diagonal scattering matrix with unit-magnitude diagonal entries, which only applies a phase shift to the incoming signal.
		The idea was later generalized to \gls{bd}-\gls{ris} with group-connected architecture \cite{Shen2020a}, where adjacent elements within the same group are connected via passive reconfigurable components\footnote{Those components can be either symmetric (e.g., capacitors and inductors) or asymmetric (e.g., ring hybrids and branch-line hybrids) \cite{Ahn2006}, resulting in symmetric and asymmetric scattering matrices, respectively.}.
		This allows wave impinging on one element to propagate within the circuit and depart partially from any element in the same group.
		It can thus manipulate both amplitude and phase of the scattered wave while remaining passive.
		Such a powerful model can be realized at reduced hardware cost using tree- and forest-connected architectures by graph theory \cite{Nerini2024}.
		\gls{bd}-\gls{ris} can also function in
		% hybrid transmitting-and-reflecting mode \cite{Li2023b} and
		multi-sector mode \cite{Li2023c} for full-space coverage and multi-user support.
		Practical challenges such as channel estimation \cite{Li2024}, mutual coupling \cite{Li2023f}, and wideband modelling \cite{Li2024a} have also been studied in recent literature.
		Its beamforming effectiveness over \gls{d}-\gls{ris} and energy efficiency over active \gls{ris} and relay systems have been proved in \gls{siso} and \gls{miso} systems \cite{Shen2020a,Nerini2023,Nerini2024,Santamaria2023,Fang2023,Zhou2023,Li2023c,Soleymani2024}.
		However, the interplay between \gls{bd}-\gls{ris} and \gls{mimo} is still in the infancy stage.
		The authors of \cite{Bartoli2023} investigated the rate-optimal joint beamforming design for a fully-connected \gls{bd}-\gls{ris}-aided \gls{mimo} system where the direct channel is negligible.
		A transmitter-side \gls{bd}-\gls{ris} was introduced to
		massive \gls{mimo} systems that exploits statistical \gls{csi} for improved spectral efficiency \cite{Mishra2024}, which again assumed negligible direct channel and fully-connected \gls{bd}-\gls{ris}.
		Received power maximization with continuous-valued and discrete-valued \gls{bd}-\gls{ris} have been tackled respectively in closed form \cite{Nerini2023} and by machine learning approach \cite{Nerini2023b}, but the corresponding single-stream transceiver is rate-suboptimal.
	% \end{subsection}

@@

		Fifth, we conduct extensive simulations to validate the analytical bounds and numerical methods.
		It is concluded that:
		\begin{itemize}
			\item \gls{bd}-\gls{ris} significantly improves the dynamic range of all channel singular values and the trade-off in manipulating them, resulting in enhanced power gain and achievable rate;
			\item A fully-connected \gls{bd}-\gls{ris} may be designed in closed-form for multi-functional optimality;
			\item The shaping advantages of \gls{bd}-\gls{ris} increase with the number of scattering elements and \gls{mimo} dimensions;
			\item \gls{bd}-\gls{ris} can activate multi-stream transmission (hence achieving the asymptotic \gls{dof}) at low transmit power thanks to its singular value shaping proficiency;
			\item The rate gap between the optimal and low-complexity beamforming designs diminishes as the \gls{ris} evolves from \gls{d} to fully-connected \gls{bd};
			\item The additional optimization cost of \gls{bd}-\gls{ris} over \gls{d}-\gls{ris} is affordable and the geodesic \gls{rcg} method is efficient on large-scale problems;
			\item The proposed \gls{ris} designs are robust to channel estimation errors and can be modified for symmetry with reasonable performance cost.
		\end{itemize}
	% \end{subsection}

@@

\begin{section}{System Model}
	% \begin{subsection}{BD-RIS}
		The \gls{bd}-\gls{ris} can be modeled as an $N_\mathrm{S}$-port network \cite{Ivrlac2010} that divides into $G$ individual groups, where group $g \in [G]$ contains $N_g$ scattering elements interconnected by real-time reconfigurable components \cite{Shen2020a} such that $N_\mathrm{S} = \sum_{g=1}^G N_g$.
		Without loss of generality we assume equal group size $N_g = L \triangleq N_\mathrm{S} / G, \ \forall g$ and no mutual coupling between elements.
		For asymmetric \gls{bd}-\gls{ris}, the overall scattering matrix is block-diagonal with unitary blocks\footnote{We assume asymmetric network by default to establish a benchmark for passive \gls{ris}. Some symmetric solutions (i.e., $\mathbf{\Theta} = \mathbf{\Theta}^\mathsf{T}$) will be discussed and their performance will be evaluated in Section~\ref{sc:simulation}.}
		\begin{equation}
			\label{eq:ris}
			\mathbf{\Theta} = \diag(\mathbf{\Theta}_1,\ldots,\mathbf{\Theta}_G),
		\end{equation}
		where $\mathbf{\Theta}_g \in \mathbb{U}^{L \times L}$ is the $g$-th diagonal block modeling the response of group $g$.
		Apparently, \gls{d}-\gls{ris} is an extreme case of \eqref{eq:ris} with group size $L=1$.
		Some viable architectures of \gls{bd}-\gls{ris} are illustrated in \cite[Fig. 3]{Shen2020a}, \cite[Fig. 5]{Li2023c}, \cite[Fig. 2]{Nerini2024} where the circuit topology have been modeled in $\mathbf{\Theta}$.
	% \end{subsection}

	% \begin{subsection}{MIMO Point-to-Point Channel}
		Consider a \gls{bd}-\gls{ris}-aided \gls{mimo} point-to-point channel with $N_\mathrm{T}$ and $N_\mathrm{R}$ transmit and receive antennas, respectively, and $N_\mathrm{S}$ scattering elements at the \gls{bd}-\gls{ris}.
		This configuration is denoted as $N_\mathrm{T} \times N_\mathrm{S} \times N_\mathrm{R}$.
		Let $\mathbf{H}_\mathrm{D} \in \mathbb{C}^{N_\mathrm{R} \times N_\mathrm{T}}$, $\mathbf{H}_\mathrm{B} \in \mathbb{C}^{N_\mathrm{R} \times N_\mathrm{S}}$, $\mathbf{H}_\mathrm{F} \in \mathbb{C}^{N_\mathrm{S} \times N_\mathrm{T}}$ denote the direct (i.e., transmitter-receiver), backward (i.e., \gls{ris}-receiver), and forward (i.e., transmitter-\gls{ris}) channels, respectively.
		The equivalent channel depends on the \gls{bd}-\gls{ris} scattering matrix
		\begin{equation}
			\label{eq:channel}
			\mathbf{H} = \mathbf{H}_\mathrm{D} + \mathbf{H}_\mathrm{B} \mathbf{\Theta} \mathbf{H}_\mathrm{F} = \mathbf{H}_\mathrm{D} + \sum_g \mathbf{H}_{\mathrm{B},g} \mathbf{\Theta}_g \mathbf{H}_{\mathrm{F},g},
			%  \triangleq \mathbf{H}_\mathrm{D} + \sum_g \mathbf{H}_g,
		\end{equation}
		where $\mathbf{H}_{\mathrm{B},g} \in \mathbb{C}^{N_\mathrm{R} \times L}$ and $\mathbf{H}_{\mathrm{F},g} \in \mathbb{C}^{L \times N_\mathrm{T}}$ are the backward and forward channels associated with group $g$, corresponding to the $(g{-}1)L{+}1$ to $gL$ columns of $\mathbf{H}_\mathrm{B}$ and rows of $\mathbf{H}_\mathrm{F}$, respectively.
		% , and $\mathbf{H}_g \triangleq \mathbf{H}_{\mathrm{B},g} \mathbf{\Theta}_g \mathbf{H}_{\mathrm{F},g}$ is the indirect channel via group $g$.
		Since unitary matrices constitute an algebraic group with respect to multiplication, the scattering matrix of group $g$ can be decomposed as
		\begin{equation}
			\label{eq:ris_decompose_group}
			\mathbf{\Theta}_g = \mathbf{L}_g \mathbf{X}_g \mathbf{R}_g^\mathsf{H},
		\end{equation}
		where $\mathbf{L}_g, \mathbf{R}_g \in \mathbb{U}^{L \times L}$ are two unitary matrices and $\mathbf{X}_g$ is a permutation matrix.
		Let $\mathbf{H}_g \triangleq \mathbf{H}_{\mathrm{B},g} \mathbf{\Theta}_g \mathbf{H}_{\mathrm{F},g}$ be the indirect channel via group $g$ and $\mathbf{H}_{\mathrm{B/F},g} = \mathbf{U}_{\mathrm{B/F},g} \mathbf{\Sigma}_{\mathrm{B/F},g} \mathbf{V}_{\mathrm{B/F},g}^\mathsf{H}$ be the \gls{svd} of the backward and forward channels, respectively.
		The equivalent channel is thus
		\begin{equation}
			\label{eq:channel_svd}
			\mathbf{H} = \overbrace{\mathbf{H}_\mathrm{D} + \sum_g \mathbf{U}_{\mathrm{B},g} \mathbf{\Sigma}_{\mathrm{B},g} \underbrace{\mathbf{V}_{\mathrm{B},g}^\mathsf{H} \mathbf{L}_g \mathbf{X}_g \mathbf{R}_g^\mathsf{H} \mathbf{U}_{\mathrm{F},g}}_\text{backward-forward} \mathbf{\Sigma}_{\mathrm{F},g} \mathbf{V}_{\mathrm{F},g}^\mathsf{H}}^\text{direct-indirect}.
		\end{equation}

		\begin{remark}
			In \eqref{eq:channel_svd}, the \gls{bd}-\gls{ris} performs a blockwise unitary transformation to combine the backward-forward (intra-group, multiplicative) channels and direct-indirect (inter-group, additive) channels.
			These two attributes are refined respectively as:
			\begin{itemize}
				\item \emph{Branch matching:} To pair and combine the branches (i.e., entries) of $\mathbf{H}_{\mathrm{B},g}$ and $\mathbf{H}_{\mathrm{F},g}$ through $\mathbf{\Theta}_g$;
				\item \emph{Mode alignment:} To align and order the modes (i.e., singular vectors) of $\{\mathbf{H}_g\}_{g \in [G]}$ with those of $\mathbf{H}_\mathrm{D}$ through $\mathbf{\Theta}$.
			\end{itemize}
		\end{remark}

		\begin{example}[\gls{siso} channel gain maximization]
			\label{eg:siso}
			Denote the direct, backward, forward channels as $h_\mathrm{D}$, $\mathbf{h}_\mathrm{B} \in \mathbb{C}^{N_\mathrm{S} \times 1}$, and $\mathbf{h}_\mathrm{F}^\mathsf{H} \in \mathbb{C}^{1 \times N_\mathrm{S}}$, respectively.
			In this case, mode alignment boils down to phase matching and the optimal \gls{bd}-\gls{ris} structure is
			% any $L \in [N_\mathrm{S}]$, including \gls{d}-\gls{ris}, suffices for perfect mode alignment using
			\begin{equation}
				\mathbf{\Theta}_{\textnormal{P-max},g}^\textnormal{SISO} = \frac{h_\mathrm{D}}{\lvert h_\mathrm{D} \rvert} \mathbf{V}_{\mathrm{B},g} \mathbf{U}_{\mathrm{F},g}^\mathsf{H}, \quad \forall g,
			\end{equation}
			where $\mathbf{V}_{\mathrm{B},g} = \bigl[\mathbf{h}_{\mathrm{B},g}/\lVert \mathbf{h}_{\mathrm{B},g} \rVert, \mathbf{N}_{\mathrm{B},g}\bigr] \in \mathbb{U}^{L \times L}$, $\mathbf{U}_{\mathrm{F},g} = \bigl[\mathbf{h}_{\mathrm{F},g}/\lVert \mathbf{h}_{\mathrm{F},g} \rVert, \mathbf{N}_{\mathrm{F},g}\bigr] \in \mathbb{U}^{L \times L}$, and $\mathbf{N}_{\mathrm{B/F},g} \in \mathbb{C}^{L \times (L-1)}$ are the orthonormal bases of kernels of $\mathbf{h}_{\mathrm{B/F},g}$.
			It is evident that any group size $L$ (including \gls{d}-\gls{ris}) suffices for perfect phase matching.
			On the other hand, the resulting maximum channel gain depends on $L$
			\begin{equation}
				\ \lvert h \rvert = \lvert h_\mathrm{D} \rvert + \sum_g \sum_l \lvert h_{\mathrm{B},g,\pi_{\mathrm{B},g}(l)} \rvert \lvert h_{\mathrm{F},g,\pi_{\mathrm{F},g}(l)} \rvert,
			\end{equation}
			where $h_{\mathrm{B/F},g,l}$ are the $l$-th entries of $\mathbf{h}_{\mathrm{B/F},g}$, and $\pi_{\mathrm{B/F},g}$ are permutations of $[L]$ sorting their magnitude in similar orders.
			That is, the maximum \gls{siso} channel gain is attained when each \gls{bd}-\gls{ris} group, apart from phase shifting, matches the $l$-th strongest backward and forward channel branches.
			A larger $L$ provides more flexible branch matching and thus higher channel gain.
		\end{example}

@@

	\begin{subsection}{Analytical Shaping Bounds}
		\label{sc:shaping_analytical}
		The main results of this subsection are presented in the following Propositions and Corollaries.
		\begin{proposition}[Degrees of freedom]
			\label{pp:dof}
			\gls{bd}-\gls{ris} may achieve a larger or smaller \gls{mimo} \gls{dof}\footnote{Here \gls{dof} refers to the maximum number of streams that can be transmitted in parallel over a \gls{mimo} channel, i.e., $\textnormal{DoF}(\mathbf{H}) = \lim_{\rho \to \infty} \frac{\log \det(\mathbf{I} + \rho \mathbf{H} \mathbf{H}^\mathsf{H})}{\log \rho}$ where $\rho$ is the \gls{snr}. It is closely related to radar \gls{dof} which refers to the number of independent measurements.} than \gls{d}-\gls{ris}.
		\end{proposition}
		\begin{proof}
			Please refer to Appendix~\ref{ap:dof}.
		\end{proof}

		While increasing the \gls{dof} improves the asymptotic rate performance for point-to-point transmission, the potential to reduce the \gls{dof} can be exploited to orthogonalize channels and align interference in multi-user scenarios.
		% reducing the \gls{dof} does not necessarily degrade the rate at finite \gls{snr}.
		% , it does not necessarily imply a higher rate at finite \gls{snr}.
		% It is worth mentioning that a larger \gls{dof} implies more parallel streams or less interference, which is beneficial for high-rate communication.
